---
title: 'ISRNM Global prevalence of PEW: Supplemental Information'
author: "(June 11, 2018)"
date: ""
output:
  html_document:
    depth: 3 # up to three depths of headings (specified by #, ## and ### etc.)
    number_sections: no # if you want number sections at each table header
    toc: yes # table of content true
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}

library(knitr)
library(htmlTable)
library(metafor)

# Read the data:
PEW <- read.csv(
  "../ProcessedData/PEW_manuscript_data_2018-05-08.csv",
  header=TRUE, sep=",", quote="\"", dec=".", na.strings = c("NA"), stringsAsFactor=FALSE)
PEW$Country <- factor(PEW$Country)
PEW$GNI_group <- factor(PEW$GNI_group,
                        levels = c("high-income country",
                                   "middle-income country",
                                   "low-income country"))
PEW$Year_group <- factor(PEW$Year_group,
                         levels = c("2000-2004","2005-2009","2010-2014"))
PEW$Study_size <-  factor(PEW$Study_size,
                          levels = c("less than 100 patients",
                                     "100-250 patients",
                                     "more than 250 patients"))
PEW$Qi_grouped <- factor(PEW$Qi_grouped,
                         levels = c("Qi below 5/8", "Qi 5/8 or higher"))
PEW$PD_HD_CKD_Tx <- factor(PEW$PD_HD_CKD_Tx)
PEW$Region_PEW <- factor(PEW$Region_PEW, # Order Region_PEW (for tables and figures)
                         levels = c(
                           "Northern America (USA)",
                           "Latin Am./Caribbean",
                           "Europe",
                           "Northern Africa (Egypt)",
                           "Sub-Saharan Africa",
                           "Western Asia",
                           "Eastern Asia",
                           "South-eastern Asia",
                           "Southern Asia",
                           "Oceania (Australia)"))

```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 80)
```


(R-execution that created this file: `r date()`.)

&nbsp;
&nbsp;

*****

**Naming conventions:** We follow the United Nation’s association of
countries with geographical regions
and refer to countries by their short forms which may or may not coincide with
the name used by that country in official documents. See the
[United Nations M49 Standard "Standard Country or Area Codes for Statistical Use"](https://unstats.un.org/unsd/methodology/m49/). For brevity
in our tables and figures we make
the following exceptions from the M49 Standard names:
“Hong Kong” refers to China, Hong Kong Special
Administrative Region; “Korea” to the Republic of Korea; “Taiwan” to the
island of Taiwan; “UK” to the United Kingdom of Great Britain and
Northern Ireland; “USA” to the United States of America;
“Iran” to Iran (Islamic Republic of).

*****
&nbsp;

*****

**Note on studies:**

*Brown EA - 2010*: This study reports from two separate patient populations, one
receiving Hemodialysis (HD) and one Peritoneal Dialysis (PD).  We refer to these
two parts as *Brown EA – 2010a for the HD-group of patients* and as 
*Brown EA - 2010b for the PD-group of patients* and count these as separate studies.

*Sanabria M - 2008*: This study reports from two separate patient populations, one
receiving Hemodialysis (HD) and one Peritoneal Dialysis (PD).  We refer to these
two parts as *Sanabria M - 2008a for the HD-group of patients* and as 
*Sanabria M - 2008b for the PD-group of patients* and count these as separate studies.

*Hecking E – 2004*: This publication reports results for participating sites from
five European countries and 
contains the needed information (Tables 1 and 4) to separate PEW prevalence (moderately and severely malnourished patients as indicated in Table 4) for all five countries. We separate Hecking E – 2004 into 5 parts/studies and refer to them as *Hecking E – 2004a for France*, *Hecking E – 2004b for Germany*, *Hecking E – 2004c for Italy*, *Hecking E – 2004d for Spain*, and *Hecking E – 2004e for UK*.  All counts of studies below are with respect to these five separate parts, i.e., Hecking E – 2004 formally contributes with five studies to Europe, of which two are in Southern Europe (Italy, Spain), two in Western Europe (France, Germany), and one in Northern Europe
(UK; the [United Nations M49 Standard](https://unstats.un.org/unsd/methodology/m49/) groups UK as Northern Europe). This assures coherent counts of studies/participants in the tables shown below. Our final meta-regression model includes region as a fixed effect and does not utilize sub-regions; we have verified that none of our conclusions change when Hecking E – 2004 is included as one study attributed to Europe (not shown).

*****
&nbsp;
&nbsp;




# 1 Distribution of study searches
<a href="#top"> [Back to top] </a>

Distribution of study searches within the IRSNM PEW prevalence consortium:

1. **Pediatric CKD**: Rulan Parekh (MD, Canada) and Sharon Russo (MD, Canada)
2. **Adult non-dialysis dependent patients with CKD stages 3-5**: Carla M Avesani (RD, Brazil)
3. **Adult dialysis-dependent patients**
    a. **In Northern America**: Adriana M Hung (MD, USA) and Rosa K Hand (RD, USA)
    b. **In Latin America and the Caribbean**: Denise Mafra (RD, Brazil), and Angeles Espinosa-Cuevas (RD, Mexico) 
    c. **In Asia**: 
        i. **East and Central Asia**: Angela YM Wang (MD, China) and Jongha Park (MD, Korea). 
        ii. **South Asia**: Anita Saxena (MD, India) 
        iii. **Southeast Asia**: Tilakavati Karupaiah (RD, Malaysia)
        iv. **Western Asia**: Siren Sezer (MD, Turkey) 
    d. **In Europe**: 
        i. **Eastern Europe**: Michal Chmielewski (MD, Poland)
        ii. **Northern Europe**: Peter Marckmann, (MD, Denmark), Lina R Johanssen (RD, UK)
        iii. **Southern Europe**: Daniel Teta (MD, Switzerland) and Lu Yimin (MD, Switzerland)
        iv. **Western Europe**: Fitsum Guebre-Egziabher (MD, France) 
    e. **In Oceania**: Maria Chan (RD, Australia) 
    f. **In Africa**: 
        i. **Southern and Eastern Africa**: Cecile Verseput (RD, South Africa)
        ii. **Western and Middle Africa**: Fatiu Arogundade (MD, Nigeria) 
        iii. **Northern Africa**: Siren Sezer (MD, Turkey)
5. **Acute-kidney injury**: Enrico Fiaccadori (MD, Italy), Talat A Ikizler (MD, USA)
6. **Kidney transplant patients**: Miklos Z Molnar (MD, USA), Pieter M Ter Wee (MD, Netherlands)

<a href="#top"> [Back to top] </a>


# 2 Quality index score
<a href="#top"> [Back to top] </a>

The quality index accounts for 5 different aspects of a study as given below.

The maximum raw score was 8 points, representing the highest methodological quality. Disagreements in the scores were resolved by discussion and consensus. The quality index was then "normalized" to the range 0 - 1 by dividing the raw score by 8 (max. achievable).



|   | Question/Property       | Answer/Feature             | Score |
|:-:|:------------------------|:---------------------------|:-----:|
| 1 | Were the target population and the observation period well defined? | - Yes<br>- No | 1<br>0 |
| 2 | Does the study discuss in detail how the SGA or MIS was done and it was performed correctly? | - Yes<br>- No | 1<br>0 |
| 3 | Method of case ascertainment | - Community survey/multiple institutions<br>- Inpatient/inpatients and outpatients/case registers<br>- Not specified| 2<br>1<br>0 |
| 4 | Administration of measurement protocol | - Administered interview/physical exam<br>- Chart diagnosis/case records<br>- Not specified | 2<br>1<br>0 |
| 5 | Catchment Area | - Prospective, broadly representative study<br>- Prospective, single center/limited catchment area<br>- Convenience sample (i.e., a cohort created for a different purpose) | 2<br>1<br>0 |
| | | *Possible points (combined):*<br>**_Range final score (points/8):_** | *0 - 8*<br>**_0 - 1_** |


<a href="#top"> [Back to top] </a>




# 3 Dialysis studies ordered according to covariates/traits

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}
mydata <- escalc(measure = "PR", xi = Cases, ni = N, data = PEW, append = TRUE)

# Random-effects model ---------------------------------------------------------
pew.res <- rma(yi, vi, data = mydata)

```
 


## 3.1 Dialysis studies ordered according to type of dialysis


```{r, echo=FALSE, message=FALSE, warning=FALSE}

pew.mes_test <- rma(yi, vi, mods = ~ PD_HD_CKD_Tx, data = mydata)

```


Observation: There does not seem to be a clear effect of the 
type of dialysis with respect to PEW prevalence in the study
(p = `r round(summary(pew.mes_test)$QMp, 3)`).

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

# Ordered according to type of dialysis:
forest(pew.res, slab = mydata$Study_name,
       order = rev(order(mydata$PD_HD_CKD_Tx, mydata$Country, (max(mydata$N) - mydata$N))),
       showweights = FALSE,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1), #refline = coef(pew.res)
       ilab = cbind(as.character(mydata$PD_HD_CKD_Tx), as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = 0.75,
       addfit = FALSE)

op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("Dialysis", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

abline(h=c(61.5+4), col = "gray")

par(op)


```
<a href="#top"> [Back to top] </a>



## 3.2 Dialysis studies ordered according to study size


```{r, echo=FALSE, message=FALSE, warning=FALSE}

pew.mes_test <- rma(yi, vi, mods = ~ N, data = mydata)
pew.mes_test_grouped <- rma(yi, vi, mods = ~ Study_size, data = mydata)

```


Observation: The size of the study does not seem to allow
conclusions about the PEW prevalence (p = `r round(summary(pew.mes_test)$QMp, 3)`;
p = `r round(summary(pew.mes_test_grouped)$QMp, 3)` for grouped study size).
But note the increasing confidence limits as study size decreases (as expected). 

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

forest(pew.res, slab = mydata$Study_name,
       order = order((max(mydata$N) - mydata$N), mydata$Country),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(as.character(mydata$Region_PEW), as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("Region", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

par(op)


```
<a href="#top"> [Back to top] </a>

## 3.3 Dialysis studies ordered according to PEW determination



```{r, echo=FALSE, message=FALSE, warning=FALSE}

pew.mes_test <- rma(yi, vi, mods = ~ PEW_determination, data = mydata, method = "ML")
# Note: For comparing models that differ in fixed effect terms (anova below),
#       maximum likelihood estimates need to be used (method = "ML";
#       default is "REML"):
#       anova(pew.mes_test3,pew.mes_test2)
pew.mes_test2 <- rma(yi, vi, mods = ~ Region_PEW, data = mydata, method = "ML")
pew.mes_test3 <- rma(yi, vi, mods = ~ PEW_determination + Region_PEW, data = mydata, method = "ML")

```


Observation: The applied method for PEW determination does not seem to allow
conclusions about the PEW prevalence reported in a study
(p = `r round(anova(pew.mes_test3,pew.mes_test2)$pval, 3)` when adjusting for region;
p = `r round(summary(pew.mes_test)$QMp, 3)` unadjusted indicates that
PEW determination is confounded with region).

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

forest(pew.res, slab = mydata$Study_name,
       order = order(mydata$PEW_determination, mydata$Country, (max(mydata$N) - mydata$N)),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(as.character(mydata$PEW_determination), as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("PEW determ.", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

abline(h=c(26.5, 68.5+4, 76.5+4), col = "gray")

par(op)


```
<a href="#top"> [Back to top] </a>

## 3.4 Dialysis studies ordered according to GNI

```{r, echo=FALSE, message=FALSE, warning=FALSE}

pew.mes_test <- rma(yi, vi, mods = ~ GNI_group, data = mydata)

```

Observation: The GNI of the country where the study is reported from does
not seem to allow conclusions about the PEW prevalence reported in a study (p = `r round(summary(pew.mes_test)$QMp, 3)`).

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

# Ordered according to GNI:

forest(pew.res, slab = mydata$Study_name,
       order = order(mydata$GNI_group, mydata$Country, (max(mydata$N) - mydata$N)),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(as.character(mydata$GNI_group), as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("GNI", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

abline(h=c(4.5, 43.5), col = "gray")

par(op)


```
<a href="#top"> [Back to top] </a>

## 3.5 Dialysis studies ordered according to Quality Index


```{r, echo=FALSE, message=FALSE, warning=FALSE}

pew.mes_test <- rma(yi, vi, mods = ~ Qi, data = mydata)
pew.mes_test_grouped <- rma(yi, vi, mods = ~ Qi_grouped, data = mydata)

```


Observation: The quality of a study as determined by our quality index (see S2 above)
does not seem to allow conclusions about the PEW prevalence reported in a study (p = `r round(summary(pew.mes_test)$QMp, 3)`; p = `r round(summary(pew.mes_test_grouped)$QMp, 3)`
for dichotomous quality index above/below 5/8).


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

# Ordered according to quality index:

forest(pew.res, slab = mydata$Study_name,
       order = order((max(mydata$Qi) - mydata$Qi), mydata$Country, (max(mydata$N) - mydata$N), decreasing = FALSE),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(formatC(mydata$Qi, format = 'f', digits = 2), as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("Qual. index", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

par(op)


```
<a href="#top"> [Back to top] </a>


## 3.6 Dialysis studies ordered according to publication year

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pew.mes_test <- rma(yi, vi, mods = ~ Year_pub, data = mydata, method = "ML")
pew.mes_test2 <- rma(yi, vi, mods = ~ Region_PEW, data = mydata, method = "ML")
pew.mes_test3 <- rma(yi, vi, mods = ~ Region_PEW + Year_pub, data = mydata, method = "ML")

```


Observation: The year of publication
does not seem to allow conclusions about the PEW prevalence reported in a study
(p = `r round(anova(pew.mes_test3,pew.mes_test2)$pval, 3)` when adjusting for region;
p = `r round(summary(pew.mes_test)$QMp, 3)` unadjusted indicates that
publication year is confounded with region).


We conclude that it is not the case that more recent studies start to agree
more on PEW prevalence.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

forest(pew.res, slab = mydata$Study_name,
       order = order( (max(mydata$Year_pub) - mydata$Year_pub),
                      mydata$Country,
                      (max(mydata$N) - mydata$N)),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(mydata$Year_pub,
                          as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("Publ. year", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

par(op)

```


<a href="#top"> [Back to top] </a>


## 3.7 Dialysis studies ordered according to PEW prevalence

Observation: This demonstrates where the name "catterpilar plot" for
these plots comes from. When ordered by prevalence, the aligned individual estimates
visually resemble a hairy caterpillar.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

forest(pew.res, slab = mydata$Study_name,
       order = order(mydata$PEW_prevalence, mydata$Country, (max(mydata$N) - mydata$N)),
       showweights = FALSE,
       psize = 1,
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(paste(formatC(100*(mydata$PEW_prevalence), format = 'f', digits = 1), "%", sep=""),
                          as.character(mydata$Country), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.65, -0.65, -0.3, -0.1),
       ilab.pos = c(2, 4, 2, 2),
       cex = .75,
       addfit = FALSE)


op <- par(cex = .75, font = 2)
text(c(-0.75, -0.58, -0.38, -0.2), 89+3, c("Prevalence", "Country", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

par(op)


```

<a href="#top"> [Back to top] </a>


# 4 Synthesis of results: Fixed-effects model/meta regression for dialysis studies

```{r, echo=FALSE, message=FALSE, warning=FALSE}

mydata <- escalc(measure = "PR", xi = Cases, ni = N, data = PEW, append = TRUE)


# Random-effects model ---------------------------------------------------------
pew.res_rand <- rma(yi, vi, data = mydata)

# mixed-effects model ---------------------------------------------------------
pew.mes <- rma(yi, vi, mods = ~ Region_PEW, data = mydata)
pew.mes_noint <- rma(yi, vi, mods = ~ Region_PEW -1, data = mydata)

table_regions <- data.frame(Region = substr(names(summary(pew.mes_noint)$b[,1]), start = 11, stop = 50),
                            Estimate = formatC((summary(pew.mes_noint)$b[,1]), format = "f", digits = 2),
                            lower95 = formatC(pmax(0, summary(pew.mes_noint)$b[,1] - 1.96*summary(pew.mes_noint)$se), format = "f", digits = 2),
                            upper95 = formatC(pmin(1, summary(pew.mes_noint)$b[,1] + 1.96*summary(pew.mes_noint)$se), format = "f", digits = 2)
                            )

colnames(table_regions) <- c("Region", "PEW-prevalence", "C.I. lower bound (95%)", "C.I. upper bound (95%)")

```

&nbsp;
&nbsp;

*****
**For comparison**: A simple random-effects model renders the following statistics
(test for heterogeneity is highly significant,
`r ifelse(summary(pew.res_rand)$QEp < 0.0001, "p < .0001 ", paste0("p = ", round(summary(pew.res_rand)$QEp, 4)))`).

| Statistic  | Explanation       | Point estimate             | 95% Conf. Int. |
|:-----:|:-----------------------------|:--------------:|:--------------:|
| $\tau^2$ | Estimated amount of residual heterogeneity | `r round(confint(pew.res_rand)$random["tau^2", "estimate"], 3)` |  (`r formatC(confint(pew.res_rand)$random["tau^2", "ci.lb"], format = 'f', digits = 3)`,  `r formatC(confint(pew.res_rand)$random["tau^2", "ci.ub"], format = 'f', digits = 3)`) |
| $I^2 (\%)$ | Ratio of residual heterogeneity and unaccounted variability | `r round(confint(pew.res_rand)$random["I^2(%)", "estimate"], 2)` | (`r round(confint(pew.res_rand)$random["I^2(%)", "ci.lb"], 2)`,  `r round(confint(pew.res_rand)$random["I^2(%)", "ci.ub"], 2)`) |


*****
&nbsp;
&nbsp;
 
 
 
**Meta-regression**: The moderator Region in this model is highly significant (
`r ifelse(summary(pew.mes)$QMp < 0.0001, "p < .0001 ", paste0("p = ", round(summary(pew.mes)$QMp, 3)))`
) and 
"explains" about `r paste0( round(summary(pew.mes)$R2, 0), "%")` of the observed heterogeneity in the data (R^2^ = `r round(summary(pew.mes)$R2, 2)`%).
The test for residual heterogeneity is highly significant, 
`r ifelse(summary(pew.mes)$QEp < 0.0001, "p < .0001 ", paste0("p = ", round(summary(pew.mes)$QEp, 4)))`).


| Statistic  | Explanation       | Point estimate             | 95% Conf. Int. |
|:-----:|:-----------------------------|:--------------:|:--------------:|
| $\tau^2$ | Estimated amount of residual heterogeneity | `r round(confint(pew.mes)$random["tau^2", "estimate"], 3)` |  (`r formatC(confint(pew.mes)$random["tau^2", "ci.lb"], format = 'f', digits = 3)`,  `r formatC(confint(pew.mes)$random["tau^2", "ci.ub"], format = 'f', digits = 3)`) |
| $I^2 (\%)$ | Ratio of residual heterogeneity and unaccounted variability | `r round(confint(pew.mes)$random["I^2(%)", "estimate"], 2)` | (`r round(confint(pew.mes)$random["I^2(%)", "ci.lb"], 2)`,  `r round(confint(pew.mes)$random["I^2(%)", "ci.ub"], 2)`) |


&nbsp;
&nbsp;


Since this is a mixed-effects model with Region as moderator, there is no longer
a single overall effect attributed to the studies.
Instead, one such level can be derived for each 
Region (level of moderator).

&nbsp;
&nbsp;

**Table**: Estimates for the average PEW prevalence in the respective regions
(model based).  *Due to the remaining high residual heterogeneity these should
be taken as broadly indicative at best.* Each study's individual random
sampling error would be in addition to this source of uncertainty.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

kable(table_regions, row.names = FALSE, format = "markdown",
      align = c("l", "c", "c", "c"))

```

&nbsp;
&nbsp;


The (model based) average PEW prevalences for each region are shown by the gray
polygons in the forest plot. Their respective width corresponds to the confidence interval
limits for the fitted values (model-based "expected" value based on the Region
where the study was conducted; each study's individual random sampling error would be
in addition to this source of uncertainty).


<a href="#top"> [Back to top] </a>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=22, fig.width=11}

forest(pew.mes, slab = mydata$Study_name,
       order = order(mydata$Region_PEW, mydata$N),
       showweights = FALSE,
       addfit = TRUE, # Adds fitted values in models with moderators.
       addcred=FALSE, # FALSE = Width of polygons reflects the confidence interval limits. 
       xlim = c(-1.5, 1.5),
       alim = c(0,1),
       clim = c(0,1),
       at = (0:5)/5,
       refline = c(0, 0.2, 0.4, 0.6, 0.8, 1),
       ilab = cbind(as.character(mydata$Region_PEW), mydata$N, mydata$Cases), 
       ilab.xpos = c(-0.45, -0.3, -0.1),
       ilab.pos = 2,
       cex = .75)

op <- par(cex = .75, font = 2)
text(c(-0.5, -0.38, -0.2), 89+3, c("Region", "N", "Cases"))
text(-1.5,           89+3, "Author(s) and Year",     pos = 4)
text(1.5,            89+3, "Prop. [95% CI]", pos = 2)

par(op)

```


The funnel plot does not suggest that publications would be systematically missing
due to publication or other reporting bias: The figure shows each study's residual
value (estimated - overall PEW proportion) against the respective estimates'
standard error (largely determined by the study size). The latter's scale is
inverted with highest precision (largest study sizes) at the top. In the absence
of publication and other reporting biases, one would expect that the dots scatter
symmetrically around the vertical line at zero residual value.  A few studies
with low precision have high positive residual values (dots on the bottom to
the right) which are not "balanced" by comparatively negative ones. Irrespective
of this minor observation, we find that the bulk of the observations is rather
balanced and therefore are not concerned with respect to publication or reporting biases.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

funnel(pew.mes, main = "Mixed-Effects Model")


```


<a href="#top"> [Back to top] </a>

# 5 Supplemental tables (dialysis studies)

For the hierarchy of geographcial regions and subregions
we follow the United Nation’s association of countries with geographical regions
and refer to countries by their short forms which may or may not coincide with
the name used by that country in official documents. See the
[United Nations M49 Standard "Standard Country or Area Codes for Statistical Use"](https://unstats.un.org/unsd/methodology/m49/) .

<a href="#top"> [Back to top] </a>

## 5.1 Contributing dialysis studies by regions as used for this study
```{r, echo=FALSE, message=FALSE, warning=FALSE}
this.text <- function(x){unique(as.character(x))}
count.names <- function(x){length(unique(x))}
my.table <- merge(aggregate(Country ~ Region_PEW, data = PEW, FUN = this.text),
                  merge(aggregate(Study_name ~ Region_PEW, data = PEW, FUN = count.names),
                        merge(aggregate(Cases ~ Region_PEW, data = PEW, FUN = sum), 
                              aggregate(N ~ Region_PEW, data = PEW, FUN = sum),
                              by = "Region_PEW"),
                        by = "Region_PEW"),
                  by = "Region_PEW")
my.table <- data.frame(my.table, Raw_percent = round(100*(my.table$Cases/my.table$N), 1))

colnames(my.table) <- c("Region (as used in this study)", "Countries","No. of studies", "Combined Cases", "Combined N", "Raw % Cases")

my.table[ , "Region (as used in this study)"] <- as.character(my.table[ , "Region (as used in this study)"])
my.table <- rbind(my.table,
                  c("Total", " ", sum(my.table[,3]), sum(my.table[,4]), sum(my.table[,5]),
                    round(100*(sum(my.table[,4])/sum(my.table[,5])), 1)))
kable(my.table, format = "markdown",
      align = c("l", "l", "r", "r", "r", "r"))

```

<a href="#top"> [Back to top] </a>

## 5.2 Contributing dialysis studies by geographic region
```{r, echo=FALSE, message=FALSE, warning=FALSE}
this.text <- function(x){unique(as.character(x))}
count.names <- function(x){length(unique(x))}
my.table <- merge(aggregate(Country ~ Geographic_Region, data = PEW, FUN = this.text),
                  merge(aggregate(Study_name ~ Geographic_Region, data = PEW, FUN = count.names),
                        merge(aggregate(Cases ~ Geographic_Region, data = PEW, FUN = sum), 
                              aggregate(N ~ Geographic_Region, data = PEW, FUN = sum),
                              by = "Geographic_Region"),
                        by = "Geographic_Region"),
                  by = "Geographic_Region")
my.table <- data.frame(my.table, Raw_percent = round(100*(my.table$Cases/my.table$N), 1))
colnames(my.table) <- c("Geographic Region", "Countries","No. of studies", "Combined Cases", "Combined N", "Raw % Cases")

my.table[ , "Geographic Region"] <- as.character(my.table[ , "Geographic Region"])
my.table <- rbind(my.table,
                  c("Total", " ", sum(my.table[,3]), sum(my.table[,4]), sum(my.table[,5]),
                    round(100*(sum(my.table[,4])/sum(my.table[,5])), 1)))
kable(my.table, format = "markdown",
      align = c("l", "l", "r", "r", "r", "r"))

```

<a href="#top"> [Back to top] </a>

## 5.3 Contributing dialysis studies by geographic subregion
```{r, echo=FALSE, message=FALSE, warning=FALSE}
this.text <- function(x){unique(as.character(x))}
count.names <- function(x){length(unique(x))}
my.table <- merge(aggregate(Country ~ Geographic_Subregion, data = PEW, FUN = this.text),
                  merge(aggregate(Study_name ~ Geographic_Subregion, data = PEW, FUN = count.names),
                        merge(aggregate(Cases ~ Geographic_Subregion, data = PEW, FUN = sum), 
                              aggregate(N ~ Geographic_Subregion, data = PEW, FUN = sum),
                              by = "Geographic_Subregion"),
                        by = "Geographic_Subregion"),
                  by = "Geographic_Subregion")
my.table <- data.frame(my.table, Raw_percent = round(100*(my.table$Cases/my.table$N), 1))
colnames(my.table) <- c("Geographic Subregion", "Countries","No. of studies", "Combined Cases", "Combined N", "Raw % Cases")

my.table[ , "Geographic Subregion"] <- as.character(my.table[ , "Geographic Subregion"])
my.table <- rbind(my.table,
                  c("Total", " ", sum(my.table[,3]), sum(my.table[,4]), sum(my.table[,5]),
                    round(100*(sum(my.table[,4])/sum(my.table[,5])), 1)))
kable(my.table, format = "markdown",
      align = c("l", "l", "r", "r", "r", "r"))

```

<a href="#top"> [Back to top] </a>

## 5.4 Contributing dialysis studies by GNI group
```{r, echo=FALSE, message=FALSE, warning=FALSE}
this.text <- function(x){unique(as.character(x))}
count.names <- function(x){length(unique(x))}
my.table <- merge(aggregate(Country ~ GNI_group, data = PEW, FUN = this.text),
                  merge(aggregate(Study_name ~ GNI_group, data = PEW, FUN = count.names),
                        merge(aggregate(Cases ~ GNI_group, data = PEW, FUN = sum), 
                              aggregate(N ~ GNI_group, data = PEW, FUN = sum),
                              by = "GNI_group"),
                        by = "GNI_group"),
                  by = "GNI_group")
my.table <- data.frame(my.table, Raw_percent = round(100*(my.table$Cases/my.table$N), 1))
colnames(my.table) <- c("GNI Group", "Countries","No. of studies", "Combined Cases", "Combined N", "Raw % Cases")

my.table[ , "GNI Group"] <- as.character(my.table[ , "GNI Group"])
my.table <- rbind(my.table,
                  c("Total", " ", sum(my.table[,3]), sum(my.table[,4]), sum(my.table[,5]),
                    round(100*(sum(my.table[,4])/sum(my.table[,5])), 1)))
kable(my.table, format = "markdown",
      align = c("l", "l", "r", "r", "r", "r"))

```

<a href="#top"> [Back to top] </a>


# Software/R packages used for this document
<a href="#top"> [Back to top] </a>

The following software/R-packages have been very helpful in compiling this document and
are gratefully acknowledged.


| Package  | Reference       |
|:-----:|:---------------------------------------------------------------------|
| R | `r citation("base")$textVersion` |
| R Studio | RStudio Team (2016). RStudio: Integrated Development for R. RStudio, Inc., Boston, MA URL http://www.rstudio.com/. |
| knitr | `r citation("knitr")$textVersion` |
| rmarkdown | `r citation("rmarkdown")$textVersion`
| metafor | `r citation("metafor")$textVersion`  |



<a href="#top"> [Back to top] </a>

<!-- # Details about R session that created this document -->
<!-- <a href="#top"> [Back to top] </a> -->

<!-- Date and time of R-execution that created this file: `r date()`. -->


<!-- ```{r, echo = FALSE} -->
<!-- devtools::session_info() -->
<!-- ``` -->


<!-- <a href="#top"> [Back to top] </a> -->


*****

End of document

*****
